---
title: "Homework_4_FINAL"
author: "Lincoln Bohn"
date: "11/4/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Homework Part 1


```{r}
rep(0, 5) # Replicate 0, 5 times
rep(1:3, 4) # Replicate 1-3, 4 times
rep(4:6, c(1, 2, 3))# Replicate 4-6, repeat 1st variable 1 time, repeat 2nd variable 2 times, repeat 3rd variable 3 times 
```

```{r}
My_Variable <- 3 # Defines My_Variable as 3
  print(My_Variable) # Print My_Variable
seq(1,10, by = 2) # Sequence 1-10, incrementally by 2
vector_1 <- c(1, 2, 3, 4, 5) # Create a vector variable vector_1, defines each row as it aligns inside c(...)
  print(vector_1) # Print vector_1
```

Homework Part 2

```{r}
data(mtcars) # Imports mtcars data
names(mtcars) # Names of mt cars variables
ncol(mtcars) # Number of columns in mtcars
nrow(mtcars) # Number of rows in mtcars
```
#3 How many rows and how many variables belong to this dataframe? What does this dataset represent and where does it come from?
    
    There are 32 rows (observations) and 11 columns (variables) | The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

#4 Descriptive statistics by function
```{r}
#Seperate functions for descriptive statistics in mtcars, mpg column

min(mtcars$mpg)
median(mtcars$mpg)
mean(mtcars$mpg)
max(mtcars$mpg)

```
#4B Summary of common descriptive statistics

```{r}
#Summary of common descriptive statistics 
summary(mtcars)
```


#5 Creates a tables showing the summary statistics of mpg by cyl.

```{r}
mtcars_df <- mtcars # mtcars_df is defined with data from mtcars for later manipulation if needed
  aggregate(mpg ~ cyl, data = mtcars_df, summary) # Computes summary of mpg by cyl, using mtcars_df, with a summary of descriptive statistics

```
```{r}
min_ag <-  # Defines variable min_ag as the minimum mpg explained by cyl.
  aggregate(mpg ~ cyl,
        data = mtcars_df, 
        FUN = min
        )
min_ag # Prints min ag

max_ag <- # Defines variable max_ag as the maximum mpg explained by cyl
  aggregate(mpg ~ cyl,
        data = mtcars_df, 
        FUN = max
        )
max_ag # Prints max ag
```

    The least fuel efficient 4 cyl car is:  Volvo 142E 21.4

    The least fuel efficient 6 cyl car is:  Merc 280C 17.8

    The least fuel efficient 8 cyl car is:  Cadillac Fleetwood & Lincoln Continental 10.4
    
    The most fuel efficient 8 cyl car is: Pontiac Firebird 19.2





#6 Create a boxplot showing MPG by cylinder
```{r}
 mpg_cyl_bplot <- # Defines variable as a boxplot of mpg by cyl, from mtcars, with some renaming and recoloring
  boxplot(mpg ~ cyl, data = mtcars, xlab = "Cylinder", ylab = "Miles Per Gallon", col = "light yellow", col.axis = "12") 

```
```{r}
library(ggplot2)
  # Testing ggplot package 
ggplot(mtcars) +
    geom_point(mapping = aes(x = cyl, y = mpg, color = cyl, size = mpg)) 
 

```

#7 Create a scatterplot of MPG versus weight
```{r}
# Scatterplot of wt and mpg, renamed axis and colors
plot(mtcars$wt, 
     mtcars$mpg, 
     xlab = "Miles Per Gallon", 
     ylab = "Weight",
     col = "Dark Green",
     col.axis = "darkred")

```
 
    In my view, Miles Per Gallon is negatively correlated with Car Weight. The strength of this relationship looks above R= -.5


#8 What is the correlation coefficient between MPG and weight? Does this confirm your answer to the previous question?
```{r}
# Prints correlation coefficient of the two variables
cor(mtcars$wt, 
    mtcars$mpg)
```

    Yes. I believed it would be over .5, it is stronger than my estimate. I was being too conservative.


#9 Create a linear regression that predicts MPG as a function of cylinders, displacement, horsepower, and weight.
```{r}
# Defines mpgreg as the linear model of mpg as a function of cyl, displacement, horsepower, and weight, from mtcars data
mpgreg <-
  lm(mpg ~ cyl + disp + hp + wt,
     data = mtcars)
# Prints summary of mpgreg variable
summary(mpgreg)
```

#10 How strong is the model overall? Which variables were significant predictors of MPG?
   
    The model is quite strong with an adjusted R-squared of 0.82, weight is a significant predictor at the strongest level possible, followed by Cylinder at the .05 level. Horsepower was not found to be a significant predictor of MPG

Homework Part 3

#11 Create a plot of the correlation matrix between all the variables in the mtcars dataset.

```{r}
library(corrplot) #loads corrplot in this chunk

mtcars_cor <- cor(mtcars_df) #defines correlation of all variables in mtcars

 corrplot(mtcars_cor, # using correlations from mtcars data, defines the graphic printed below.
          method = 'number', # Shows correlation coefficient strength in graphic
          type = "upper", #
          order = "hclust") 
  # Another correlation matrix graphic from researching ??corrplot
 
 corrplot(mtcars_cor,
        p.mat = ,
        method = 'circle',
        type = 'lower',
        insig='blank',
        addCoef.col ='black',
        number.cex = 0.8,
        order = 'AOE',
        diag=FALSE)

```

#11B What does this uncover about the correlations between the explanatory variables in the model you constructed for #9?
    
      This graphic uncovers the strength of the correlation we did from #9, If we go to a signifcant correlation like mpg ~ cyl, we can see -.85, which is a strong relationship. I would suspect this. Lets see mpg ~ wt, -.87, a little stronger, but also expected. Finally, mpg ~ hp also shows a -.78 strong negative relationship, but from #9 p=.102 which is telling us that there is a large chance that gross horsepower is not a good explanatory for mpg, because it could be due to another, possibly random, factor

#12 Re-create the regression that you ran for #6 in the SPSS homework assignment within R and interpret the results.

        
```{r}
library(foreign) #library "foreign" package


df_spss_BB <- # Defines df_spss_BB as reading an SPSS file, and using its original labels imported into the data frame
  read.spss("CPUC_Broadband_CTs.sav", use.value.label=TRUE, to.data.frame=TRUE) 


```

    #12B Construct another multivariate regression that predicts Adoption Rate, this time including at least five independent variables, two of which should be ethnic/racial variables. How strong is the model overall? Which variables are statistically significant predictors?
      
```{r}
spss_BB_lm <- # Defines variable as linear model of adoption rate as explained by other variables
  lm(AdoptionRate ~ medage + MHI + White + Hispanic + pct25wbach, data = df_spss_BB)
 # Summary of desctriptive statistics for the linear model
summary(spss_BB_lm)

```
    
    Adjusted R-squared values are the same across spss and R calculations. spss = .433, .4329 in the model above, so even further decimal points, though they are not necessary. Therefore, the model is medium strength, showing our calculations are correct across two programs! 
    Nearly All variables are significant at the strongest level, this was true in both programs. In both programs, medage was slightly less significant, but it is negligible at p = .002, whereas in this program it is p=0.00221. 
    
